---
title: "Office of National Statistics (UK) Baby Names"
#author: "Jeremy Aytoun"
format: html
server: shiny
---

## Why did I build this?

I wanted to investigate names and when they were first and latest recorded as births after seeing a statistic for names in a trivia game "Shot in the Dark". While seeing Quarto and Shiny project. I took the top 100 names from the listings and some names outside of the top 100 as was curious for those names.

So if were wanting to use the data for top 100, would have to filter the data set

In process of creating this, I discovered there is  similar tool created by [Anna Powell Smith](https://names.darkgreener.com/). Though, I am using this as a learning experience on R Shiny and Quarto.

Data set Columns: Rank, Year, Name, Gender, Count of births

```{r Loading Data}
#| include: false
#| context: data
library(tidyverse)
library(ggplot2)
library(shiny)
names = readxl::read_xlsx("names.xlsx",sheet="names")
names$Names <- str_to_title(names$Names)

```

Total number of births are counted from 1996 on wards as the data set from 1904-1994 did not include number of births.

```{r}
library(ggplot2)


selectInput("names_options","Names:", unique(sort(names$Names)))
plotOutput("outPlot")

textOutput("textOutPut")


```

Lower rank shows more babies born with this designated name during that year.

```{r}
#| context: server



library(tidyverse)
library(ggthemes)
library(ggrepel)


output$outPlot <- renderPlot({
  filtered_data <- reactive({
  dplyr::filter(names, Names == input$names_options)
    

})
 
  
ggplot(filtered_data(), aes(x = Year, y = RANK, label = paste0("Year: ",Year, "\nRank: ", RANK)))+ geom_point(color="#FF00CC") + geom_line(color = "#CD5C5C") + geom_area(fill = "#FF6F00", alpha = 0.5) +geom_label_repel() +
  labs(title = paste0(str_to_title(filtered_data()$Names), " Total: ", sum(filtered_data()$Count)), 
       subtitle = paste0("First recorded in year: ",  min(filtered_data()$Year),  "\nLatest recording in: ", max(filtered_data()$Year)))+theme_fivethirtyeight() 


})


```

# About the data:

The data is from Office of National Statistics, merged two data sets one from [1904 - 1994](https://www.ons.gov.uk/peoplepopulationandcommunity/birthsdeathsandmarriages/livebirths/datasets/babynamesenglandandwalestop100babynameshistoricaldata) which was set for every 10 years and another from [1996-2020](https://www.ons.gov.uk/peoplepopulationandcommunity/birthsdeathsandmarriages/livebirths/datasets/babynamesinenglandandwalesfrom1996) which is yearly data.

All data belongs to [Office for National Statistics](https://www.ons.gov.uk/peoplepopulationandcommunity/birthsdeathsandmarriages/livebirths/bulletins/babynamesenglandandwales/2020/relateddata)

Data is Crown Copyright, reproduced under [Open Government License](https://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/)

I merged the data into one file to be used
